---
title: "FRP surpisal"
author: "Rosy Southwell"
subtitle: "FRP estimated using deconvolution in MNE"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
knit: (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                        encoding = encoding, 
                        output_file = "FRP_LMM_mne_allfix.html"
                        )})
---

```{r}
knitr::opts_chunk$set(cache=F)
USE_DC <- T
version <- "FRP_GLM_allfix_decim10"
description <- "using MNE, all fixations, removing outlier fixations"
knitname<- paste0("FRP_LMM_mne_",version,"_", (if(USE_DC){"dc"} else {"nodc"}), ".html")
```
# Info
### USE_DC: `r USE_DC`
### version: `r version`
### description: `r description`

# Data
```{r, message=FALSE, warning=FALSE, cache=T}
# List of libraries to be installed and loaded
libraries <- c(
  "lme4", "lmerTest", "reshape2", "plyr", "beepr", "car", "sjPlot", "emmeans",
  "ggplot2", "tidyr", "readxl", "optimx", "dfoptim", "robustlmm", "mediation",
  "stringr", "ggpubr", "gridExtra", "lattice", "blme", "DescTools", "dplyr", "tidyverse","magrittr"
)

# Function to install and load libraries
install_and_load <- function(libraries) {
  for (lib in libraries) {
    if (!require(lib, character.only = TRUE)) {
      install.packages(lib, dependencies = TRUE)
      library(lib, character.only = TRUE)
    }
  }
}

# Install and load the libraries
install_and_load(libraries)
options(scipen = 50)
source("rutils.R") # colScale
```

load data

```{r, message=F, warning=F}
text_file = '/Users/roso8920/Emotive Computing Dropbox/Rosy Southwell/EML Rosy/EMLLM/info/ia_label_mapping_opt_surprisal.csv'

# eeg_file = '/Volumes/Blue1TB/EEG_processed/FRP_GLM_allfix_decim10/FRP_stats/ALL_dcFRP_stats_formatted.csv' 
eeg_file = '/Volumes/Blue1TB/EEG_processed/FRP_GLM_allfix_decim10/FRP_stats/ALL_dcFRP_stats.csv'

# df_orig <- read.csv(eeg_file) 
```

subjects excluded 
```{r, message=F, warning=F}
# # using QC based on reading v sham classifier <.55 AUROC
# exclude <-c()
# # exclude <- c("EML1_034", "EML1_055", "EML1_093", "EML1_099" ,"EML1_115", "EML1_132", "EML1_154")
# # using MW rates - between 20 and 80% 
# # using comprehension not significantly different from chance
# df <- df %>% filter(!ParticipantID %in% exclude)
```

## Merge with text info and filter fixations
```{r}
df<-df_orig
if( 'EVENT' %in% colnames(df)){df <-df%>% rename(identifier=EVENT)}
# DATA TYPES
LOGICAL_VARS <- c('punctuation','stop_word')
df[LOGICAL_VARS] <- sapply(df[LOGICAL_VARS],as.logical)
# delta_wf = log(min(df$word_freq[df$word_freq>0]))-1 # use the minimum word frequency for any missing word frerquencies so we can take log
fixfillval = mean(df$fixDur, na.rm=T)
df <- df %>% rename(fixDur=duration_sec)
df <- df %>% mutate(
                  logWordFreq = ifelse(is.na(log_word_freq),0,log_word_freq),
                  logFixDur = ifelse(is.na(fixDur), 0, log(fixDur))
)
# replace RMS_all with RMSall to make later string munging easier
colnames(df) <- gsub("RMS_all", "RMSall", colnames(df))
colnames(df) <- gsub("_lat","lat", colnames(df))
colnames(df) <- gsub("max_abs","maxabs", colnames(df))


```



# transforms 
```{r}

# refixation
df <- df %>% group_by(ParticipantID, identifier, IA_ID) %>% mutate(
  IA_fix_count = n(),
  refixation_ = ifelse(n()>1,1,0))
# remove some unused cols as this is getting ungainly
df <- df %>% select(c(-contains(c("AFF","CCP","PPO9","PPO10","FCz","N400_PPO","1_CPz","N400_CPz_max","N1_*_max","P1_*_min","max_abs","zero_crossings")) , contains("RMS")))

feats_eye <- grep("fix|pupil|duration",colnames(df), value=T, ignore.case=T)
feats_eeg <- grep("N400|P1|N1",colnames(df), value=T, ignore.case=T)
feats_lex = c("punctuation","sentence_ix", "word_in_sentence", "sentence_word_count", "relative_word_position", "stop_word","surprisal")
feats <- c(feats_eye, feats_eeg)

# # mutate so all predictor cols are scaled 
# # scale_this <- function(x){
# #   (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
# # }
# ftarray = array(df %>% ungroup() %>% select(feats) %>% unlist(), dim=c(nrow(df), length(feats)), dimnames =list(NULL, feats))
# ftarray <- colScale(ftarray)
# df[,feats] <- ftarray
# df <- df %>% mutate(across(feats,colScale(.)))

refixations <- df %>% group_by(ParticipantID) %>% summarise(refix_rate = mean(refixation_, na.rm=T))
# df <- df %>% filter(refixation_==0)

df_sub <- df %>% group_by(ParticipantID) %>% summarise_if(is.numeric,mean, na.rm=T)
# drop rows with no IA
df.words <- drop_na(df, "IA_LABEL") %>% filter(punctuation==F) %>% filter(stop_word==F)
df_sub_words <- df.words %>% group_by(ParticipantID) %>% summarise_if(is.numeric,mean, na.rm=T)

```

## predict fixation fixDur from EEG features
### for all fixations and for content word fixaitons
```{r, message=F, warning=F}
m.eye <- lmer(fixDur ~ (1|identifier) + (1|ParticipantID) + 
                P1_PPOave_mean + P1_PPOave_max + P1_PPOave_maxlat + 
                N1_PPOave_mean + N1_PPOave_min + N1_PPOave_minlat + 
                N400_CPz_mean+ N400_CPz_min + N400_CPz_minlat,
              data=df)
m.eye.words <- lmer(fixDur ~ (1|identifier) + (1|ParticipantID) + 
                P1_PPOave_mean + P1_PPOave_max + P1_PPOave_maxlat + 
                N1_PPOave_mean + N1_PPOave_min + N1_PPOave_minlat + 
                N400_CPz_mean+ N400_CPz_min + N400_CPz_minlat,
              data=df.words)
tab_model(m.eye, m.eye.words)
```


# Modeling with comprehension/MW
## Load Behavioural data

```{r, message=F, warning=F}

df_comp <- read.csv('/Users/roso8920/Emotive Computing Dropbox/Rosy Southwell/EyeMindLink/Processed/Behaviour/EML1_page_level.csv')
df_comp <- df_comp %>% rename(ParticipantID=ParticipantID) %>% 
  mutate(Page=PageNum-1,
         identifier=paste0(Text, Page),
         meanX = rowMeans(select(., Rote_X, SVT, Inference_X, Deep_X), na.rm = TRUE),
         meanY = rowMeans(select(., Rote_Y, Inference_Y), na.rm = TRUE), 
         meanZ = rowMeans(select(., Rote_Z, Inference_Z, Deep_Z), na.rm = TRUE)
  )
df <- merge(df, df_comp, on=c("ParticipantID","identifier"))
df.words <- merge(df.words, df_comp, on=c("ParticipantID","identifier"))
```

# Eye-mind-linkage metric at page level
note atanh transform for correlation when it is to be used in model
```{r, message=F, warning=F}
eml.page.words <- df.words %>% group_by(ParticipantID, identifier) %>% summarise(
    n_wordfix=n(),
    cor_N400meanCPz_surprisal = cor(surprisal, N400_CPz_mean),
    cor_N400meanCPz_logWordFreq = cor(logWordFreq, N400_CPz_mean),
    cor_N400minCPz_surprisal = cor(surprisal, N400_CPz_min),
    cor_N400minCPz_logWordFreq = cor(logWordFreq, N400_CPz_min),
    cor_N400minlatCPz_surprisal = cor(surprisal, N400_CPz_minlat),
    cor_N400minlatCPz_logWordFreq = cor(logWordFreq, N400_CPz_minlat),
    cor_P1_surprisal = cor(surprisal, P1_PPOave_mean),
    cor_N1_surprisal = cor(surprisal, N1_PPOave_mean),
    cor_P1_logWordFreq = cor(logWordFreq, P1_PPOave_mean),
    cor_N1_logWordFreq = cor(logWordFreq, N1_PPOave_mean),
    cor_logFixDur_surprisal = cor(surprisal, logFixDur),
    # fixdur
    logWordFixDur = mean(logFixDur, na.rm=TRUE),
    cor_word_P1mean_logFixDur = cor(P1_PPOave_mean,logFixDur),
    cor_word_P1max_logFixDur = cor(P1_PPOave_max,logFixDur),
    cor_word_P1lat_logFixDur = cor(P1_PPOave_maxlat,logFixDur),
    cor_word_N1mean_logFixDur = cor(N1_PPOave_mean,logFixDur),
    cor_word_N1min_logFixDur = cor(N1_PPOave_min,logFixDur),
    cor_word_N1lat_logFixDur = cor(N1_PPOave_minlat,logFixDur),
    cor_word_N400mean_logFixDur = cor(N400_CPz_mean,logFixDur),
    cor_word_N400min_logFixDur = cor(N400_CPz_min,logFixDur),
    cor_word_N400lat_logFixDur = cor(N400_CPz_minlat,logFixDur),
    # 
    meansurprisal=mean(surprisal, na.rm=TRUE),
    meanP1word=mean(P1_PPOave_mean, na.rm=TRUE),
    meanN1word=mean(N1_PPOave_mean, na.rm=TRUE),
    maxP1word = mean(P1_PPOave_max, na.rm=TRUE),
    latP1word = mean(P1_PPOave_maxlat, na.rm=TRUE),
    minN1word = mean(N1_PPOave_min, na.rm=TRUE),
    latN1word= mean(N1_PPOave_minlat, na.rm=TRUE),
    meanN400word=mean(N400_CPz_mean, na.rm=TRUE),
    minN400word=mean(N400_CPz_min, na.rm=TRUE),
    latN400word=mean(N400_CPz_minlat, na.rm=TRUE)
) %>%  ungroup() %>% 
  filter(n_wordfix>2) # remove instances with single or two fixation on page as correlation coeff will be 1 or -1
# eml.page.words[Reduce(`&`, lapply(eml.page.words, is.finite),]

eml.page.all <- df %>% dplyr::group_by(ParticipantID, identifier) %>% dplyr::summarise(
    n_fix=n(),
    meanP1=mean(P1_PPOave_mean, na.rm=TRUE),
    meanN1=mean(N1_PPOave_mean, na.rm=TRUE),
    maxP1 = mean(P1_PPOave_max, na.rm=TRUE),
    latP1 = mean(P1_PPOave_maxlat, na.rm=TRUE),
    minN1 = mean(N1_PPOave_min, na.rm=TRUE),
    latN1= mean(N1_PPOave_minlat, na.rm=TRUE),
    meanN400=mean(N400_CPz_mean, na.rm=TRUE),
    minN400=mean(N400_CPz_min, na.rm=TRUE),
    latN400=mean(N400_CPz_minlat, na.rm=TRUE),
    cor_P1mean_logFixDur = cor(P1_PPOave_mean,logFixDur,use="pairwise.complete.obs" ),
    cor_P1max_logFixDur = cor(P1_PPOave_max,logFixDur,use="pairwise.complete.obs"),
    cor_P1lat_logFixDur = cor(P1_PPOave_maxlat,logFixDur,use='pairwise.complete.obs'),
    cor_N1mean_logFixDur = cor(N1_PPOave_mean,logFixDur,use='pairwise.complete.obs'),
    cor_N1min_logFixDur = cor(N1_PPOave_min,logFixDur,use='pairwise.complete.obs'),
    cor_N1lat_logFixDur = cor(N1_PPOave_minlat,logFixDur,use='pairwise.complete.obs'),
    cor_N400mean_logFixDur = cor(N400_CPz_mean,logFixDur,use='pairwise.complete.obs'),
    cor_N400min_logFixDur = cor(N400_CPz_min,logFixDur,use='pairwise.complete.obs'),
    cor_N400lat_logFixDur = cor(N400_CPz_minlat,logFixDur,use='pairwise.complete.obs'),
    logFixDur = mean(logFixDur, na.rm=TRUE)

) %>%  ungroup() %>% 
  filter(n_fix>2) # remove instances with single or two fixation on page as correlation coeff will be 1 or -1
# eml.page.all[Reduce(`&`, lapply(eml.page.all, is.finite)),]

df_page <- left_join(eml.page.words, df_comp, by=c('ParticipantID', 'identifier')) 
df_page <- left_join(df_page, eml.page.all, by=c('ParticipantID', 'identifier')) 
df_page$MW = as.factor(df_page$MW)
df_page$identifier =as.factor(df_page$identifier)
df_page$ParticipantID = as.factor(as.character(df_page$ParticipantID ))
write_csv(df_page, 'page_lvl_linkage_feats.csv')

# z score the new metrics
feats_page <- grep("cor_|mean|lat|min|max|fix",colnames(df_page), value=T)
ftarray = array(df_page %>% ungroup() %>% select(feats_page) %>% unlist(), dim=c(nrow(df_page), length(feats_page)), dimnames =list(NULL, feats_page))
ftarray <- colScale(ftarray)
df_pageZ <- df_page
df_pageZ[,feats_page] <- ftarray
write_csv(df_pageZ, 'page_lvl_linkage_feats_Z.csv')
# truncate the z scores to 3 sd
df_pageZtrunc <- df_pageZ
df_pageZtrunc[,feats_page] <- pmin(pmax(ftarray, -3), 3)
write_csv(df_pageZtrunc, 'page_lvl_linkage_feats_Z_trunc.csv')

```

## plot the new metrics
```{r, message=F, warning=F}

```

## binomial regression: predict page level MW from page level EEG
```{r, message=F, warning=F}

m.mw.main.word <- glmer(MW ~ meanP1word + meanN1word + meanN400word + minN400word+latN400word+minN1word +maxP1word +latN1word+latP1word+ logFixDur +
                      n_wordfix + logFixDur + (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
tab_model(m.mw.main.word )

```
## binomial regression: predict page level outcome from page level correlations with logFixDur, all words
### MW
```{r, message=F, warning=F}
m.mw.fixdur <- glmer(MW ~ 
                    cor_word_P1mean_logFixDur + cor_word_P1max_logFixDur + cor_word_P1lat_logFixDur + 
                    cor_word_N1mean_logFixDur + cor_word_N1min_logFixDur + cor_word_N1lat_logFixDur + 
                    cor_word_N400mean_logFixDur + cor_word_N400min_logFixDur + cor_word_N400lat_logFixDur + 
                     n_wordfix + (1|identifier) + (1|ParticipantID), data=df_pageZ %>% drop_na(MW), family = "binomial")
tab_model(m.mw.fixdur)
```
### Comp, word fix
```{r, message=F, warning=F}
comp_label <-'Rote_X'
form <- as.formula(paste0(comp_label, "~", 
                    "cor_word_P1mean_logFixDur + cor_word_P1max_logFixDur + cor_word_P1lat_logFixDur + 
                    cor_word_N1mean_logFixDur + cor_word_N1min_logFixDur + cor_word_N1lat_logFixDur + 
                    cor_word_N400mean_logFixDur + cor_word_N400min_logFixDur + cor_word_N400lat_logFixDur + n_wordfix + (1|identifier) + (1|ParticipantID)"))
m.comp.wfixdur <- glmer(form, data=df_pageZ %>% drop_na(comp_label), family = "binomial")
tab_model(m.comp.wfixdur)
```
### Comp, all fix
```{r, message=F, warning=F}
comp_label <-'Rote_X'
form <- as.formula(paste0(comp_label, "~", 
                    "cor_P1mean_logFixDur + cor_P1max_logFixDur + cor_P1lat_logFixDur + 
                    cor_N1mean_logFixDur + cor_N1min_logFixDur + cor_N1lat_logFixDur + 
                    cor_N400mean_logFixDur + cor_N400min_logFixDur + cor_N400lat_logFixDur + n_fix + (1|identifier) + (1|ParticipantID)"))
m.comp.fixdur <- glmer(form, data=df_page %>% drop_na(comp_label), family = "binomial")
tab_model(m.comp.fixdur)
```
### Mean_X
```{r, message=F, warning=F}
m.meanx.fixdur <- lmer(meanX ~ 
                    cor_word_P1mean_logFixDur + cor_word_P1max_logFixDur + cor_word_P1lat_logFixDur + 
                    cor_word_N1mean_logFixDur + cor_word_N1min_logFixDur + cor_word_N1lat_logFixDur + 
                    cor_word_N400mean_logFixDur + cor_word_N400min_logFixDur + cor_word_N400lat_logFixDur + 
                    n_wordfix +  (1|identifier) + (1|ParticipantID), data=df_pageZ %>% drop_na(meanX))
tab_model(m.meanx.fixdur)
```
## binomial regression: predict page level MW from page level EEG feats, all fixations
```{r, message=F, warning=F}

m.mw.main <- glmer(MW ~ meanP1 + meanN1 + meanN400 + minN400+latN400+minN1 +maxP1 +latN1+latP1+
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
# tab_model(m.mw.mean, m.mw.min, m.mw.max)
# tab_model(m.mw.mean)
tab_model(m.mw.main,m.mw.main.word)

```
## binomial regression: predict page level comprehensionfrom page level correlations with surprisal
```{r, message=F, warning=F}
comp_label <-'Rote_X'
m.comp.all <- glmer(Rote_X ~ 
                    cor_P1_logWordFreq + cor_P1_surprisal +  cor_N1_logWordFreq  + cor_N1_surprisal +
                    cor_N400meanCPz_logWordFreq + cor_N400meanCPz_surprisal   +
                    meanP1 + meanN1 + meanN400 +
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(Rote_X), family = "binomial")
m.comp.N400 <- glmer(Rote_X ~ 
                    cor_N400meanCPz_logWordFreq + cor_N400meanCPz_surprisal   + 
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(Rote_X), family = "binomial")

# tab_model(m.comp.mean, m.comp.min, m.comp.max)
# tab_model(m.comp.mean)
tab_model(m.comp.all)
tab_model(m.comp.N400)

```

# COMPREHENSION
## binomial regression: predict page level comp from page level correlations with surprisal, all fix
```{r, message=F, warning=F}

m.comp.main <- glmer(Rote_X ~ meanP1 + meanN1 + meanN400 + minN400+latN400+minN1 +maxP1 +latN1+latP1+
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(Rote_X), family = "binomial")
# tab_model(m.mw.mean, m.mw.min, m.mw.max)
# tab_model(m.mw.mean)
tab_model(m.comp.main)

```
## Does page-average surprisal predict MW or comprehension
```{r}
m.mw.sp <- glmer(MW ~ meansurprisal  +  (1+meansurprisal|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
tab_model(m.mw.sp)

m.comp.sp <- glmer(Rote_X ~ meansurprisal  +  (1+meansurprisal|ParticipantID), data=df_page %>% drop_na(Rote_X), family = "binomial")
tab_model(m.comp.sp)
```

