---
title: "FRP surpisal"
author: "Rosy Southwell"
subtitle: "Fixed bug with baseline correction in EEG processing"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
knit: (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                        encoding = encoding, 
                        output_file = "FRP_LMM_v7.1_nodc.html"
                        )})
---

```{r}
USE_DC <- T
version <- "FRP_GLM_simple_decim10"
description <- "using MNE, only content words, removing outlier fixations"
knitname<- paste0("FRP_LMM_mne_",version,"_", (if(USE_DC){"dc"} else {"nodc"}), ".html")
```
# Info
## USE_DC: `r USE_DC`
## version: `r version`
## description: `r description`

# Data
```{r, message=FALSE, warning=FALSE}
# List of libraries to be installed and loaded
libraries <- c(
  "lme4", "lmerTest", "reshape2", "plyr", "beepr", "car", "sjPlot", "emmeans",
  "ggplot2", "tidyr", "readxl", "optimx", "dfoptim", "robustlmm", "mediation",
  "stringr", "ggpubr", "gridExtra", "lattice", "blme", "DescTools", "dplyr", "tidyverse","magrittr"
)

# Function to install and load libraries
install_and_load <- function(libraries) {
  for (lib in libraries) {
    if (!require(lib, character.only = TRUE)) {
      install.packages(lib, dependencies = TRUE)
      library(lib, character.only = TRUE)
    }
  }
}

# Install and load the libraries
install_and_load(libraries)
options(scipen = 50)

```

load data

```{r, message=F, warning=F}
text_file = '/Users/roso8920/Emotive Computing Dropbox/Rosy Southwell/EML Rosy/EMLLM/info/ia_label_mapping_opt_surprisal.csv'

eeg_file = '/Volumes/Blue1TB/EEG_processed/FRP_GLM_simple_decim10/FRP_stats/ALL_dcFRP_stats.csv'
df_orig <- read.csv(eeg_file) 
```

subjects excluded 
```{r, message=F, warning=F}
# # using QC based on reading v sham classifier <.55 AUROC
# exclude <-c()
# # exclude <- c("EML1_034", "EML1_055", "EML1_093", "EML1_099" ,"EML1_115", "EML1_132", "EML1_154")
# # using MW rates - between 20 and 80% 
# # using comprehension not significantly different from chance
# df <- df %>% filter(!ParticipantID %in% exclude)
```

## Merge with text info and filter fixations
```{r}
df<-df_orig
# DATA TYPES
LOGICAL_VARS <- c('punctuation','stop_word')
df[LOGICAL_VARS] <- sapply(df[LOGICAL_VARS],as.logical)
delta_wf = log(min(df$word_freq[df$word_freq>0]))-1 # use the minimum word frequency for any missing word frerquencies so we can take log
df <- df %>% rename(fixDur=duration_sec)
df <- df %>% mutate(
                  logSurprisal = log(surprisal),
                  logWordFreq = ifelse(word_freq>0,log(word_freq),delta_wf),
                  logFixDur = log(fixDur)
)
# replace RMS_all with RMSall to make later string munging easier
colnames(df) <- gsub("RMS_all", "RMSall", colnames(df))
colnames(df) <- gsub("_lat","lat", colnames(df))
colnames(df) <- gsub("max_abs","maxabs", colnames(df))
# drop rows with no IA
df <- drop_na(df, "IA_LABEL")
# drop rows with punc
df <- df %>% filter(punctuation==F)
# drop stopwords
df <- df %>% filter(stop_word==F)



```



# transforms 
```{r}
# lagged surprisal
df <- df %>% group_by(ParticipantID, identifier) %>% arrange(fix_pageIndex, .by_group=T) %>% 
  mutate(prev_Surprisal = dplyr::lag(surprisal, n=1, default=NA),
         prev_logSurprisal = dplyr::lag(logSurprisal, n=1, default=NA),
         )
# refixation
df <- df %>% group_by(ParticipantID, identifier, IA_ID) %>% mutate(
  IA_fix_count = n(),
  refixation_ = ifelse(n()>1,1,0))
df$refixation = as.factor(df$refixation_)
# remove some unused cols as this is getting ungainly
df <- df %>% select(-contains(c("AFF","CCP","PPO9","PPO10","FCz","N400_PPO","1_CPz","N400_CPz_max","N1_*_max","P1_*_min","max_abs","zero_crossings")))

feats_eye <- grep("fix|surp|freq",colnames(df), value=T)
feats_eeg <- grep("N400|P1|N1",colnames(df), value=T)
feats <- c(feats_eye, feats_eeg)
refixations <- df %>% group_by(ParticipantID) %>% summarise(refix_rate = mean(refixation_, na.rm=T))
# df <- df %>% filter(refixation_==0)

df_sub <- df %>% group_by(ParticipantID) %>% summarise_if(is.numeric,mean, na.rm=T)
```

# plots
```{r, message=F, warning=F}
df_long <- df %>% select(-refixation) %>% pivot_longer(all_of(feats), names_to="feature",values_to="value")
df_long <- df_long %>% mutate(channel=str_match(feature, '(?<component>N400|N1|P1)_(?<channel>[[:alnum:]]{2,6})_(?<feature_type>[a-zA-Z_]*)')$channel,
                              feature_type=str_match(feature, '(?<component>N400|N1|P1)_(?<channel>[[:alnum:]]{2,6})_(?<featuretype>[a-zA-Z_]*)')$featuretype,
                              eeg_component=str_match(feature, '(?<component>N400|N1|P1)_(?<channel>[[:alnum:]]{2,6})_(?<featuretype>[a-zA-Z_]*)')$component)
# compute lower and upper whiskers for each group
ylims <- df_long %>%
  group_by(feature_type) %>%
  summarise(Q1 = quantile(value, 1/100), Q3 = quantile(value, 99/100)) %>%
  ungroup()

p1<-ggplot(df_long %>% filter(feature_type=='mean')) + 
  geom_density(aes(x=value, fill=channel), alpha=0.3) + xlim(-1000,1000) + ggtitle('N400_mean')
p2<-ggplot(df_long %>% filter(grep(feature_type,'lat'))) + 
  geom_density(aes(x=value, fill=channel), alpha=0.3) + xlim(-1000,1000) + ggtitle('N400_min')
p3<-ggplot(df_long %>% filter(feature_type %in% c('min','max','maxabs')) )+ 
  geom_density(aes(x=value, fill=channel), alpha=0.3) + xlim(-1000,1000) + ggtitle('N400_max')
grid.arrange(p1,p2,p3)

ggplot(df)+
    geom_density(aes(x=fixDur))+
  scale_x_log10()

ggplot(df)+
    geom_density(aes(x=surprisal))+
  scale_x_log10()

ggplot(df_long)+
    geom_violin(aes(x=value,y=factor(channel), fill=factor(channel),outliers = FALSE), alpha=0.3)+
  facet_wrap(~feature_type, scales="free") 

# surprisal scatterplots
ggplot(df,aes(x=logSurprisal, y=logFixDur)) + 
  geom_point(size=1, alpha=0.01)+stat_cor(method="pearson")
ggplot(df_long %>% filter(feature_type =='N400_mean'),aes(x=logSurprisal,y=value, color=factor(channel)))+
    geom_point(alpha=0.01, size=0.5)+stat_cor(method="pearson", size=2, r.accuracy=0.01, p.accuracy=0.001, label.x=-7.5, label.y=1000)+
  facet_wrap(~channel, scales="free") 

```



# LME stats
## predict EEG from LOG surprisal and fixDur
```{r, message=F, warning=F}
m.l.m <- lmer(N400_CPz_mean ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal, data=df)
m.l.min <- lmer(N400_CPz_min ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal, data=df)
m.l.lat <- lmer(N400_CPz_minlat ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal, data=df)
# m.l.P1 <-lmer(P1_PPOave_mean ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal, data=df)
# m.l.N1 <- lmer(N1_PPOave_mean ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal, data=df)
tab_model(m.l.m, m.l.min , m.l.lat, show.ci=F)

```

## predict fixDur from surprisal
```{r, message=F, warning=F}
m.d <- lmer(fixDur ~ (1|identifier) + (1|ParticipantID) + logSurprisal, data=df)
tab_model(m.d)
```

## predict fixation fixDur from EEG features
```{r, message=F, warning=F}
m.eye <- lmer(fixDur ~ (1|identifier) + (1|ParticipantID) + N400_CPz_mean+ P1_PPOave_mean +N1_PPOave_mean, data=df)
tab_model(m.eye)
```

## predict EEG and surprisal from previous IA surprisal
```{r, message=F, warning=F}
mp.m <- lmer(N400_CPz_mean ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal + prev_logSurprisal, data=df)
tab_model(mp.m)
mp.s <- lm( logSurprisal ~ prev_logSurprisal, data=df)
tab_model(mp.s)
```
## model including bool for refixation
```{r, message=F, warning=F}
mr.m <- lmer(N400_CPz_mean ~ (1|identifier) + (1|ParticipantID) + logFixDur + refixation_*logSurprisal, data=df)
tab_model(mr.m)
Anova(mr.m)
plot_model(mr.m,type = "emm", terms=c("logSurprisal", "refixation_"))
tb<-tableone::CreateTableOne(data = df, vars = c("N400_CPz_mean", "logFixDur"), strata = 'refixation', test=T)
knitr::kable(print(tb))
```



## predict EEG from surprisal and other lexical variables 
```{r, message=F, warning=F}
m.lx.m <- lmer(N400_CPz_mean ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal + logWordFreq + relative_word_position, data=df )
m.lx.P1 <-lmer(P1_PPOave_mean ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal+ logWordFreq + relative_word_position, data=df)
m.lx.N1 <- lmer(N1_PPOave_mean ~ (1|identifier) + (1|ParticipantID) + logFixDur + logSurprisal+ logWordFreq + relative_word_position, data=df)
tab_model(m.lx.m, m.lx.P1, m.lx.N1, show.ci=F)

```

## predict EEG from surprisal and fixDur with random slope of surprisal for participant
```{r, message=F, warning=F}
m.rs.m <- lmer(N400_CPz_mean ~ (1+logSurprisal|ParticipantID) + (1|identifier) + logSurprisal + logFixDur+ logWordFreq + relative_word_position, data=df)
m.rs.min <- lmer(N400_max__CPz ~ (1+logSurprisal|ParticipantID) +  (1|identifier) + logFixDur + logSurprisal+ logWordFreq + relative_word_position, data=df)
m.rs.max <- lmer(N400_min_magnitude_CPz ~(1+logSurprisal|ParticipantID) +  (1|identifier) + logFixDur + logSurprisal+ logWordFreq + relative_word_position, data=df)
tab_model(m.rs.m, m.rs.min, m.rs.max, show.ci=F)
dotplot(ranef(m.rs.m))
```

# Modeling with comprehension/MW
## Load Behavioural data
```{r, message=F, warning=F}

df_comp <- read.csv('/Users/roso8920/Emotive Computing Dropbox/Rosy Southwell/EyeMindLink/Processed/Behaviour/EML1_page_level.csv')
df_comp <- df_comp %>% rename(ParticipantID=ParticipantID) %>% 
  mutate(Page=PageNum-1,
         identifier=paste0(Text, Page))
df <- merge(df, df_comp, on=c("ParticipantID","identifier"))
df_mw <- df %>% drop_na(MW) %>% mutate(MW=as.factor(MW))

```

# bobby's models: EEG ~ MW*lang at fixation level
## N400
```{r, message=F, warning=F}
# mB <- allFit(lmer(N400_CPz_mean ~ 1 + MW*logSurprisal + (MW*logSurprisal|ParticipantID) + (logSurprisal|ParticipantID:identifier), data=df_mw))
mB <- lmer(N400_CPz_mean ~ 1 + MW*logSurprisal + MW*logWordFreq +(MW*logSurprisal|ParticipantID) + +(MW*logWordFreq|ParticipantID) + (logSurprisal|ParticipantID:identifier) + (logWordFreq|ParticipantID:identifier) , data=df_mw, control=lmerControl(optimizer='Nelder_Mead'))
summary(mB)
tab_model(mB)
emmeans(mB, pairwise ~ MW, p.adjust = "fdr")
```
## P1
```{r, message=F, warning=F}
# mB <- allFit(lmer(N400_CPz_mean ~ 1 + MW*logSurprisal + (MW*logSurprisal|ParticipantID) + (logSurprisal|ParticipantID:identifier), data=df_mw))
mB <- lmer(P1_PPOave_mean ~ 1 + MW*logSurprisal + MW*logWordFreq + MW*relative_word_position +(MW*logSurprisal|ParticipantID) + +(MW*logWordFreq|ParticipantID) + (logSurprisal|ParticipantID:identifier) + (logWordFreq|ParticipantID:identifier) + (relative_word_position|ParticipantID:identifier), data=df_mw, control=lmerControl(optimizer='Nelder_Mead'))
summary(mB)
tab_model(mB)
emmeans(mB, pairwise ~ MW, p.adjust = "fdr")
```
## N1
```{r, message=F, warning=F}
# mB <- allFit(lmer(N400_CPz_mean ~ 1 + MW*logSurprisal + (MW*logSurprisal|ParticipantID) + (logSurprisal|ParticipantID:identifier), data=df_mw))
mB <- lmer(N1_PPOave_mean ~ 1 + MW*logSurprisal + MW*logWordFreq + MW*relative_word_position +(MW*logSurprisal|ParticipantID) + +(MW*logWordFreq|ParticipantID) + (logSurprisal|ParticipantID:identifier) + (logWordFreq|ParticipantID:identifier) + (relative_word_position|ParticipantID:identifier), data=df_mw, control=lmerControl(optimizer='Nelder_Mead'))
summary(mB)
tab_model(mB)
emmeans(mB, pairwise ~ MW, p.adjust = "fdr")
```

# Eye-mind-linkage metric at page level
note atanh transform for correlation when it is to be used in model
```{r, message=F, warning=F}
eml_page <- df %>% group_by(ParticipantID, identifier) %>% summarise(
    count=n(),
    cor_N400meanCPz_logSurprisal = atanh(cor(logSurprisal, N400_CPz_mean)),
    cor_N400minCPz_logSurprisal = atanh(cor(logSurprisal, N400_min_magnitude_CPz)),
    cor_N400maxCPz_logSurprisal = atanh(cor(logSurprisal, N400_max__CPz)),
    cor_N400meanCPz_logWordFreq = atanh(cor(logWordFreq, N400_CPz_mean)),
    cor_P1_logSurprisal = atanh(cor(logSurprisal, P1_PPOave_mean)),
    cor_N1_logSurprisal = atanh(cor(logSurprisal, N1_PPOave_mean)),
    cor_P1_logWordFreq = atanh(cor(logWordFreq, P1_PPOave_mean)),
    cor_N1_logWordFreq = atanh(cor(logWordFreq, N1_PPOave_mean)),
    cor_logFixDur_logSurprisal = atanh(cor(logSurprisal, logFixDur)),
    meanLogSurprisal=mean(logSurprisal),
    meanP1=mean(P1_PPOave_mean),
    meanN1=mean(N1_PPOave_mean),
    meanN400=mean(N400_CPz_mean)
) %>% drop_na() %>% 
  filter(count>2) # remove instances with single or two fixation on page as correlation coeff will be 1 or -1
eml_page[Reduce(`&`, lapply(eml_page, is.finite)),]

df_page <- left_join(eml_page, df_comp, by=c('ParticipantID', 'identifier')) 
df_page$MW = as.factor(df_page$MW)


# plot the new metrics
P1<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_N400meanCPz_logSurprisal, color=MW))
p2<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_logFixDur_logSurprisal, color=MW))
p3<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_P1_logSurprisal , color=MW))
p4<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_N1_logSurprisal , color=MW))
p5<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_P1_logWordFreq , color=MW))
p6<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_N1_logWordFreq , color=MW))
grid.arrange(P1,p2,p3,p4, p5,p6)

P1<-ggplot(df_page %>% drop_na(MW))+
  geom_boxplot( aes(x=meanP1, color=MW))
p2<-ggplot(df_page %>% drop_na(MW))+
  geom_boxplot( aes(x=meanN1, color=MW))
p3<-ggplot(df_page %>% drop_na(MW))+
  geom_boxplot( aes(x=meanN400 , color=MW))
grid.arrange(P1,p2,p3)
# plot by participantID
# ggplot(df_page, aes(x=cor_N400meanCPz_logSurprisal, y=ParticipantID)) +
#   geom_boxplot(  horizontal = TRUE,        outlier.colour="red",
#         outlier.fill="red",
#         outlier.size=0.2)

```

## binomial regression: predict page level MW from page level correlations with surprisal
```{r, message=F, warning=F}
m.mw.all <- glmer(MW ~ 
                    cor_P1_logWordFreq + cor_P1_logSurprisal +  cor_N1_logWordFreq  + cor_N1_logSurprisal +
                    cor_N400meanCPz_logWordFreq + cor_N400meanCPz_logSurprisal   + 
                    meanP1 + meanN1 + meanN400 +
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
m.mw.N400 <- glmer(MW ~ 
                    cor_N400meanCPz_logWordFreq + cor_N400meanCPz_logSurprisal   + 
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
m.mw.P1<- glmer(MW ~ 
                    cor_P1_logWordFreq + cor_P1_logSurprisal   + 
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
m.mw.N1<- glmer(MW ~ 
                    cor_N1_logWordFreq + cor_N1_logSurprisal   + 
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
m.mw.main <- glmer(MW ~ meanP1 + meanN1 + meanN400 +
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
# tab_model(m.mw.mean, m.mw.min, m.mw.max)
# tab_model(m.mw.mean)
tab_model(m.mw.all)

tab_model(m.mw.main)
tab_model(m.mw.P1)
tab_model(m.mw.N1)
tab_model(m.mw.N400)

```
## binomial regression: predict page level comprehensionfrom page level correlations with surprisal
```{r, message=F, warning=F}
comp_label <-'Rote_X'
m.comp.all <- glmer(Rote_X ~ 
                    cor_P1_logWordFreq + cor_P1_logSurprisal +  cor_N1_logWordFreq  + cor_N1_logSurprisal +
                    cor_N400meanCPz_logWordFreq + cor_N400meanCPz_logSurprisal   +
                    meanP1 + meanN1 + meanN400 +
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(Rote_X), family = "binomial")
m.comp.N400 <- glmer(Rote_X ~ 
                    cor_N400meanCPz_logWordFreq + cor_N400meanCPz_logSurprisal   + 
                    (1|identifier) + (1|ParticipantID), data=df_page %>% drop_na(Rote_X), family = "binomial")

# tab_model(m.comp.mean, m.comp.min, m.comp.max)
# tab_model(m.comp.mean)
tab_model(m.comp.all)
tab_model(m.comp.N400)

```

# Does page-average surprisal predict MW or comprehension
```{r}
m.mw.sp <- glmer(MW ~ meanLogSurprisal  +  (1+meanLogSurprisal|ParticipantID), data=df_page %>% drop_na(MW), family = "binomial")
tab_model(m.mw.sp)

m.comp.sp <- glmer(Rote_X ~ meanLogSurprisal  +  (1+meanLogSurprisal|ParticipantID), data=df_page %>% drop_na(Rote_X), family = "binomial")
tab_model(m.comp.sp)
```

# check crrelation between MW and comprehension
```{r}
cor.test(df_comp$MW, df_comp$Rote_X, method="spearman")

# subj average comp df
df_sub_comp <- df_comp %>% group_by(ParticipantID) %>% summarise_if(is.numeric,mean, na.rm=T)

cor.test(df_sub_comp$MW, df_sub_comp$Rote_X, method="pearson")****
```