---
title: "FRP surpisal"
author: "Rosy Southwell"
subtitle: "Fixed bug with baseline correction in EEG processing"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
knit: (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                        encoding = encoding, 
                        output_file = "FRP_LMM_v7.1_nodc.html"
                        )})
---

```{r}
USE_DC <- F
version <- "v7.1"
description <- "v7 but with 7 subjects excluded using QC based on reading v sham classifier"
knitname<- paste0("FRP_LMM_",version,"_", (if(USE_DC){"dc"} else {"nodc"}), ".html")
```
# Info
## USE_DC: `r USE_DC`
## version: `r version`
## description: `r description`

# Data
```{r, message=FALSE, warning=FALSE}
# List of libraries to be installed and loaded
libraries <- c(
  "lme4", "lmerTest", "reshape2", "plyr", "beepr", "car", "sjPlot", "emmeans",
  "ggplot2", "tidyr", "readxl", "optimx", "dfoptim", "robustlmm", "mediation",
  "stringr", "ggpubr", "gridExtra", "lattice", "blme", "DescTools", "dplyr", "tidyverse","magrittr"
)

# Function to install and load libraries
install_and_load <- function(libraries) {
  for (lib in libraries) {
    if (!require(lib, character.only = TRUE)) {
      install.packages(lib, dependencies = TRUE)
      library(lib, character.only = TRUE)
    }
  }
}

# Install and load the libraries
install_and_load(libraries)
options(scipen = 50)

```

load data

```{r, message=F, warning=F}
text_file = '/Users/roso8920/Emotive Computing Dropbox/Rosy Southwell/EML Rosy/EMLLM/info/ia_label_mapping_opt_surprisal.csv'
if (USE_DC){
  eeg_dir = '/Users/roso8920/Emotive Computing Dropbox/Rosy Southwell/EML Rosy/Data/EEG_processed/unfolded_FRP_reparsed_v7/n400_stats/'
} else {
  eeg_dir = '/Users/roso8920/Emotive Computing Dropbox/Rosy Southwell/EML Rosy/Data/EEG_processed/unfolded_FRP_reparsed_v7/n400_stats_nodc/'
}
file_pat = '_reading_N400_stats\\.csv$'
files=list.files(path = eeg_dir, pattern= file_pat)
```
Loop over participants
```{r, message=F, warning=F}
all<-c()
i<-0
for (f in files){
    i=i+1
    df <-read.csv(paste0(eeg_dir,f))
    p <- strsplit(f,'_')
    pID <- paste(p[[1]][1],p[[1]][2],sep='_')
    df$pID <- pID
    # df<- df %>% dplyr::select(c("pID", "type","latency","fixDur","task","identifier" , "page_fixation_ix","IA_ID","n400_magnitude_CPz","n400_latency_CPz"  ))
    df<- df %>% filter(task %in% c('reading','sham')) %>% droplevels()
    all[[i]] <- df
}
df_all <- do.call(rbind, all)
df<-df_all
```

subjects excluded using QC based on reading v sham classifier <.55 AUROC
```{r, message=F, warning=F}

exclude <- c("EML1_034", "EML1_055", "EML1_093", "EML1_099" ,"EML1_115", "EML1_132", "EML1_154")
df <- df %>% filter(!pID %in% exclude)
```

## Merge with text info and filter fixations
```{r}
text <- read.csv(text_file, colClasses=c("punctuation"="logical","stop_word"="logical","text"="factor","identifier"="factor"))
df<-df %>% filter(task=='reading')
delta_wf = log(min(text$word_freq[text$word_freq>0]))-1 # use the minimum word frequency for any missing word frerquencies so we can take log

text <- text %>% rename(surprisalText=opt.125m_surprisal_wholetext, surprisalPage = opt.125m_surprisal_page)

text<- text %>%mutate(logSurprisalText = log(surprisalText),
                  logSurprisalPage = log(surprisalPage),
                  logWordFreq = ifelse(word_freq>0,log(word_freq),delta_wf)
   )
df <- df %>% merge(text, on=c("IA_ID","identifier")) %>% rename(fixDur=duration)


# drop rows with no IA
df <- drop_na(df, "IA_LABEL")
# drop rows with punc
df <- df %>% filter(punctuation==F)
# drop stopwords
df <- df %>% filter(stop_word==F)

feats <- grep("n400|p1|n1|fixDur",colnames(df), value=T)
feats_eeg <- grep("n400|p1|n1",colnames(df), value=T)

```

# basic checks on text variables
```{r}
ggplot(text,aes(x=relative_word_position, y=logSurprisalText)) + 
geom_point(alpha=0.5, size=0.5)+stat_cor(method="pearson", size=4, r.accuracy=0.01, p.accuracy=0.001)
```

# transforms 
```{r}
# df<-df %>% mutate(across(contains('n400_magnitude'),
#              .fns=~log(-.x)), .names="log_neg_{.col}")
df<-df %>% mutate(
                  logSurprisalText = log(surprisalText),
                  logSurprisalPage = log(surprisalPage),
                  logWordFreq = ifelse(word_freq>0,log(word_freq),-18.98),
                  logFixDur = log(fixDur)
)
# lagged surprisal
df <- df %>% group_by(pID, identifier) %>% arrange(page_fixation_ix, .by_group=T) %>% 
  mutate(prev_surprisalText = dplyr::lag(surprisalText, n=1, default=NA),
         prev_surprisalPage = dplyr::lag(surprisalPage, n=1, default=NA),
         prev_logSurprisalText = dplyr::lag(logSurprisalText, n=1, default=NA),
         prev_logSurprisalPage = dplyr::lag(logSurprisalPage, n=1, default=NA),
         )
# refixation
df <- df %>% group_by(pID, identifier, IA_ID) %>% mutate(
  IA_fix_count = n(),
  refixation_ = ifelse(n()>1,1,0),
  refixation=as.factor(ifelse(n()>1,1,0)))

df <- df %>% filter(refixation==0)

# channel averags
df <- df %>% mutate(
  p1_mean = (p1_mean_PPO9h + p1_mean_PPO10h / 2),
  n1_mean = (n1_mean_PPO9h + n1_mean_PPO10h / 2)
)

```
# examine
```{r, message=F, warning=F}
df_sub <- df %>% group_by(pID) %>% summarise_if(is.numeric,mean, na.rm=T)
```
# plots
```{r, message=F, warning=F}
df_long <- df %>% pivot_longer(all_of(c(feats,'logFixDur')), names_to="feature",values_to="value")
df_long <- df_long %>% mutate(channel=str_match(feature, 'n400_\\w*_([[:alnum:]]{2,5})$')[,2],
                              feature_type=str_match(feature, '(n400_\\w*)_[[:alnum:]]{2,5}$')[,2])
# compute lower and upper whiskers for each group
ylims <- df_long %>%
  group_by(feature_type) %>%
  summarise(Q1 = quantile(value, 1/100), Q3 = quantile(value, 99/100)) %>%
  ungroup()

p1<-ggplot(df_long %>% filter(feature_type=='n400_mean')) + 
  geom_density(aes(x=value, fill=channel), alpha=0.3) + xlim(-1000,1000) + ggtitle('n400_mean')
p2<-ggplot(df_long %>% filter(feature_type=='n400_min_magnitude')) + 
  geom_density(aes(x=value, fill=channel), alpha=0.3) + xlim(-1000,1000) + ggtitle('n400_min')
p3<-ggplot(df_long %>% filter(feature_type=='n400_max_magnitude')) + 
  geom_density(aes(x=value, fill=channel), alpha=0.3) + xlim(-1000,1000) + ggtitle('n400_max')
grid.arrange(p1,p2,p3)

ggplot(df)+
    geom_density(aes(x=fixDur))+
  scale_x_log10()

ggplot(df)+
    geom_density(aes(x=surprisalText))+
  scale_x_log10()

ggplot(df_long)+
    geom_violin(aes(x=value,y=factor(channel), fill=factor(channel),outliers = FALSE), alpha=0.3)+
  facet_wrap(~feature_type, scales="free") 

# surprisal scatterplots
ggplot(df,aes(x=logSurprisalText, y=logFixDur)) + 
  geom_point(size=1, alpha=0.01)+stat_cor(method="pearson")
ggplot(df_long %>% filter(feature_type =='n400_mean'),aes(x=logSurprisalText,y=value, color=factor(channel)))+
    geom_point(alpha=0.01, size=0.5)+stat_cor(method="pearson", size=2, r.accuracy=0.01, p.accuracy=0.001, label.x=-7.5, label.y=1000)+
  facet_wrap(~channel, scales="free") 

```



# LME stats
## predict EEG from LOG surprisal and fixDur
```{r, message=F, warning=F}
m.l.m <- lmer(n400_mean_CPz ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText, data=df)
m.l.min <- lmer(n400_max_magnitude_CPz ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText, data=df)
m.l.max <- lmer(n400_min_magnitude_CPz ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText, data=df)
m.l.zc <- lmer(n400_zero_crossings_CPz ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText, data=df)
m.l.p1 <-lmer(p1_mean ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText, data=df)
m.l.n1 <- lmer(n1_mean ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText, data=df)
tab_model(m.l.m, m.l.p1, m.l.n1, show.ci=F)

```

## predict fixDur from surprisal
```{r, message=F, warning=F}
m.d <- lmer(fixDur ~ (1|identifier) + (1|pID) + logSurprisalText, data=df)
tab_model(m.d)
```

## predict fixation fixDur from EEG features
```{r, message=F, warning=F}
m.eye <- lmer(fixDur ~ (1|identifier) + (1|pID) + n400_mean_CPz+ p1_mean +n1_mean, data=df)
tab_model(m.eye)
```

## predict EEG and surprisal from previous IA surprisal
```{r, message=F, warning=F}
mp.m <- lmer(n400_mean_CPz ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText + prev_logSurprisalText, data=df)
tab_model(mp.m)
mp.s <- lm( logSurprisalText ~ prev_logSurprisalText, data=df)
tab_model(mp.s)
```
<!-- ## model including bool for refixation  -->
<!-- ```{r, message=F, warning=F} -->
<!-- mr.m <- lmer(n400_mean_CPz ~ (1|identifier) + (1|pID) + logFixDur + refixation*logSurprisalText, data=df) -->
<!-- tab_model(mr.m) -->
<!-- Anova(mr.m) -->
<!-- plot_model(mr.m,type = "emm", terms=c("logSurprisalText", "refixation")) -->
<!-- tb<-tableone::CreateTableOne(data = df, vars = c("n400_mean_CPz", "logFixDur"), strata = 'refixation', test=T) -->
<!-- knitr::kable(print(tb)) -->
<!-- ``` -->



## predict EEG from surprisal and other lexical variables 
```{r, message=F, warning=F}
m.lx.m <- lmer(n400_mean_CPz ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText + logWordFreq + relative_word_position, data=df )
m.lx.p1 <-lmer(p1_mean ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText+ logWordFreq + relative_word_position, data=df)
m.lx.n1 <- lmer(n1_mean ~ (1|identifier) + (1|pID) + logFixDur + logSurprisalText+ logWordFreq + relative_word_position, data=df)
tab_model(m.lx.m, m.lx.p1, m.lx.n1, show.ci=F)

```

## predict EEG from surprisal and fixDur with random slope of surprisal for participant
```{r, message=F, warning=F}
m.rs.m <- lmer(n400_mean_CPz ~ (1+logSurprisalText|pID) + (1|identifier) + logSurprisalText + logFixDur+ logWordFreq + relative_word_position, data=df)
m.rs.min <- lmer(n400_max_magnitude_CPz ~ (1+logSurprisalText|pID) +  (1|identifier) + logFixDur + logSurprisalText+ logWordFreq + relative_word_position, data=df)
m.rs.max <- lmer(n400_min_magnitude_CPz ~(1+logSurprisalText|pID) +  (1|identifier) + logFixDur + logSurprisalText+ logWordFreq + relative_word_position, data=df)
tab_model(m.rs.m, m.rs.min, m.rs.max, show.ci=F)
dotplot(ranef(m.rs.m))
```

# Modeling with comprehension/MW
## Load Behavioural data
```{r, message=F, warning=F}

df_comp <- read.csv('/Users/roso8920/Emotive Computing Dropbox/Rosy Southwell/EyeMindLink/Processed/Behaviour/EML1_page_level.csv')
df_comp <- df_comp %>% rename(pID=ParticipantID) %>% 
  mutate(Page=PageNum-1,
         identifier=paste0(Text, Page))
df <- merge(df, df_comp, on=c("pID","identifier"))
df_mw <- df %>% drop_na(MW) %>% mutate(MW=as.factor(MW))

```

# bobby's models: EEG ~ MW*lang at fixation level
## N400
```{r, message=F, warning=F}
# mB <- allFit(lmer(n400_mean_CPz ~ 1 + MW*logSurprisalText + (MW*logSurprisalText|pID) + (logSurprisalText|pID:identifier), data=df_mw))
mB <- lmer(n400_mean_CPz ~ 1 + MW*logSurprisalText + MW*logWordFreq + MW*relative_word_position +(MW*logSurprisalText|pID) + +(MW*logWordFreq|pID) + (logSurprisalText|pID:identifier) + (logWordFreq|pID:identifier) + (relative_word_position|pID:identifier), data=df_mw, control=lmerControl(optimizer='Nelder_Mead'))
summary(mB)
tab_model(mB)
emmeans(mB, pairwise ~ MW, p.adjust = "fdr")
```
## P1
```{r, message=F, warning=F}
# mB <- allFit(lmer(n400_mean_CPz ~ 1 + MW*logSurprisalText + (MW*logSurprisalText|pID) + (logSurprisalText|pID:identifier), data=df_mw))
mB <- lmer(p1_mean ~ 1 + MW*logSurprisalText + MW*logWordFreq + MW*relative_word_position +(MW*logSurprisalText|pID) + +(MW*logWordFreq|pID) + (logSurprisalText|pID:identifier) + (logWordFreq|pID:identifier) + (relative_word_position|pID:identifier), data=df_mw, control=lmerControl(optimizer='Nelder_Mead'))
summary(mB)
tab_model(mB)
emmeans(mB, pairwise ~ MW, p.adjust = "fdr")
```
## N1
```{r, message=F, warning=F}
# mB <- allFit(lmer(n400_mean_CPz ~ 1 + MW*logSurprisalText + (MW*logSurprisalText|pID) + (logSurprisalText|pID:identifier), data=df_mw))
mB <- lmer(n1_mean ~ 1 + MW*logSurprisalText + MW*logWordFreq + MW*relative_word_position +(MW*logSurprisalText|pID) + +(MW*logWordFreq|pID) + (logSurprisalText|pID:identifier) + (logWordFreq|pID:identifier) + (relative_word_position|pID:identifier), data=df_mw, control=lmerControl(optimizer='Nelder_Mead'))
summary(mB)
tab_model(mB)
emmeans(mB, pairwise ~ MW, p.adjust = "fdr")
```

# Eye-mind-linkage metric at page level
note atanh transform for correlation when it is to be used in model
```{r, message=F, warning=F}
eml_page <- df %>% group_by(pID, identifier) %>% summarise(
    count=n(),
    cor_n400meanCPz_logSurprisalText = atanh(cor(logSurprisalText, n400_mean_CPz)),
    cor_n400minCPz_logSurprisalText = atanh(cor(logSurprisalText, n400_min_magnitude_CPz)),
    cor_n400maxCPz_logSurprisalText = atanh(cor(logSurprisalText, n400_max_magnitude_CPz)),
    cor_n400meanCPz_logWordFreq = atanh(cor(logWordFreq, n400_mean_CPz)),
    cor_P1_logSurprisalText = atanh(cor(logSurprisalText, p1_mean)),
    cor_N1_logSurprisalText = atanh(cor(logSurprisalText, n1_mean)),
    cor_P1_logWordFreq = atanh(cor(logWordFreq, p1_mean)),
    cor_N1_logWordFreq = atanh(cor(logWordFreq, n1_mean)),
    cor_logFixDur_logSurprisalText = atanh(cor(logSurprisalText, logFixDur)),
    meanLogSurprisalText=mean(logSurprisalText),
    meanP1=mean(p1_mean),
    meanN1=mean(n1_mean),
    meanN400=mean(n400_mean_CPz)
) %>% drop_na() %>% 
  filter(count>2) # remove instances with single or two fixation on page as correlation coeff will be 1 or -1
eml_page[Reduce(`&`, lapply(eml_page, is.finite)),]

df_page <- left_join(eml_page, df_comp, by=c('pID', 'identifier')) 
df_page$MW = as.factor(df_page$MW)


# plot the new metrics
p1<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_n400meanCPz_logSurprisalText, color=MW))
p2<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_logFixDur_logSurprisalText, color=MW))
p3<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_P1_logSurprisalText , color=MW))
p4<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_N1_logSurprisalText , color=MW))
p5<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_P1_logWordFreq , color=MW))
p6<-ggplot(df_page %>% drop_na(MW))+
  geom_density( aes(x=cor_N1_logWordFreq , color=MW))
grid.arrange(p1,p2,p3,p4, p5,p6)

p1<-ggplot(df_page %>% drop_na(MW))+
  geom_boxplot( aes(x=meanP1, color=MW))
p2<-ggplot(df_page %>% drop_na(MW))+
  geom_boxplot( aes(x=meanN1, color=MW))
p3<-ggplot(df_page %>% drop_na(MW))+
  geom_boxplot( aes(x=meanN400 , color=MW))
grid.arrange(p1,p2,p3)
# plot by participantID
# ggplot(df_page, aes(x=cor_n400meanCPz_logSurprisalText, y=pID)) +
#   geom_boxplot(  horizontal = TRUE,        outlier.colour="red",
#         outlier.fill="red",
#         outlier.size=0.2)

```

## binomial regression: predict page level MW from page level correlations with surprisal
```{r, message=F, warning=F}
m.mw.all <- glmer(MW ~ 
                    cor_P1_logWordFreq + cor_P1_logSurprisalText +  cor_N1_logWordFreq  + cor_N1_logSurprisalText +
                    cor_n400meanCPz_logWordFreq + cor_n400meanCPz_logSurprisalText   + 
                    meanP1 + meanN1 + meanN400 +
                    (1|identifier) + (1|pID), data=df_page %>% drop_na(MW), family = "binomial")
m.mw.n400 <- glmer(MW ~ 
                    cor_n400meanCPz_logWordFreq + cor_n400meanCPz_logSurprisalText   + 
                    (1|identifier) + (1|pID), data=df_page %>% drop_na(MW), family = "binomial")
m.mw.p1<- glmer(MW ~ 
                    cor_P1_logWordFreq + cor_P1_logSurprisalText   + 
                    (1|identifier) + (1|pID), data=df_page %>% drop_na(MW), family = "binomial")
m.mw.n1<- glmer(MW ~ 
                    cor_N1_logWordFreq + cor_N1_logSurprisalText   + 
                    (1|identifier) + (1|pID), data=df_page %>% drop_na(MW), family = "binomial")
m.mw.main <- glmer(MW ~ meanP1 + meanN1 + meanN400 +
                    (1|identifier) + (1|pID), data=df_page %>% drop_na(MW), family = "binomial")
# tab_model(m.mw.mean, m.mw.min, m.mw.max)
# tab_model(m.mw.mean)
tab_model(m.mw.all)

tab_model(m.mw.main)
tab_model(m.mw.p1)
tab_model(m.mw.n1)
tab_model(m.mw.n400)

```
## binomial regression: predict page level comprehensionfrom page level correlations with surprisal
```{r, message=F, warning=F}
comp_label <-'Rote_X'
m.comp.all <- glmer(Rote_X ~ 
                    cor_P1_logWordFreq + cor_P1_logSurprisalText +  cor_N1_logWordFreq  + cor_N1_logSurprisalText +
                    cor_n400meanCPz_logWordFreq + cor_n400meanCPz_logSurprisalText   +
                    meanP1 + meanN1 + meanN400 +
                    (1|identifier) + (1|pID), data=df_page %>% drop_na(Rote_X), family = "binomial")
m.comp.n400 <- glmer(Rote_X ~ 
                    cor_n400meanCPz_logWordFreq + cor_n400meanCPz_logSurprisalText   + 
                    (1|identifier) + (1|pID), data=df_page %>% drop_na(Rote_X), family = "binomial")

# tab_model(m.comp.mean, m.comp.min, m.comp.max)
# tab_model(m.comp.mean)
tab_model(m.comp.all)
tab_model(m.comp.n400)

```

# Does page-average surprisal predict MW or comprehension
```{r}
m.mw.sp <- glmer(MW ~ meanLogSurprisalText  +  (1+meanLogSurprisalText|pID), data=df_page %>% drop_na(MW), family = "binomial")
tab_model(m.mw.sp)

m.comp.sp <- glmer(Rote_X ~ meanLogSurprisalText  +  (1+meanLogSurprisalText|pID), data=df_page %>% drop_na(Rote_X), family = "binomial")
tab_model(m.comp.sp)
```

# check crrelation between MW and comprehension
```{r}
cor.test(df_comp$MW, df_comp$Rote_X, method="spearman")

# subj average comp df
df_sub_comp <- df_comp %>% group_by(pID) %>% summarise_if(is.numeric,mean, na.rm=T)

cor.test(df_sub_comp$MW, df_sub_comp$Rote_X, method="pearson")****
```